---
title: "Exploration"
author: "Maximiliano Gonzalez"
date: "07-04-2021"
output: pdf_document
---

```{r setup2, include=T}
library(tm)
library(topicmodels)
library(wordcloud2)
library(broom)
library(tidytext)
library(caret)
knitr::opts_chunk$set(echo = F)
source('C:\\repos\\learn-doing\\R\\functions.R')

slice.projects=df%>%group_by(Codigo)%>%filter(winner=='Seleccionada')%>%ungroup%>%mutate(percentile=ntile(montoOferta,100))%>%mutate()
rm(df)
```

## Existing Classifications of Types of Project

I first examine the types of projects I have in the sample. Around 30\% are infraestructure projects. While almost 18% corresponds to residential. I examine more in detail the three types of projects. 
In regardsm to size, the average of the winning bid for comercial and indistrial projects is higher, because of a concentration of projects of higher size above the mean. However, overall there are no big differences between the distributions. 

The classification of generic product is a hierarchical breakdown of the previous classifications into more detailed classifications, for the case of comertial, which is only coded as civil works.The ONU product classification is redundant, it is a different coding for the same variable. 

It is not easy to be thorough in regards to the types of projects there are in the sample. 

```{r cars}
#Table of first classification

#Table of first and second classification 
```

I find that the detailed classification is not very reliable. First, the "housing" category includes cultural patrimony, cesfam, schools, and sports, so it is not very useful to filter by this item.The same happens with the "building"category, as buildings are sometimes sporting places. what can be said is that housing projects are almost always found under the correct category but the category is filled with other types of projects. This problem is true in general for all cateogries, they do not provide a true picture of what kinds of projects are in it, except for the "Street improvement".  The most clean classification is the one regarding streets. 

Overall, we can say that housing projects and streets are almost certainly found in the respective classifications. However, only in the street classification we can be sure of what types of projects we are dealing with. In conclusion, the thematic classification of the projects is of little value to select types of projects. 

Another interesting take is which kinds of institutions auction projects. However, this requires considerable cleaning since there are no simple identifiers. 

## Exploring Types of Buyers
here I study which types of government units develop projects. I rely in a  user-defined classification of the government units. The next table displays the types of project by government unit. Municipalities develop around 80\% of the projects in the sample.

Municipality projects ar around 140MM CLP, around 40\% of the mean in the total sample. The next important group are ministers, with 10\% opf the projects but almost 40\%of the total amount. An important buyer in the sample is the Child School Borad(JUNJI) which develops on itself around 5\% of the total amount of construction projects. The number of offerers does not differ significantly across the different units.  


```{r organisms, echo=FALSE}
organisms.table=slice.projects%>%group_by(TypeOrganism)%>%count()%>%arrange(-n)%>%ungroup()%>%mutate(prc=round(n/sum(n),2))

organisms.table.detailed=slice.projects%>%group_by(TypeOrganism)%>%summarise(n=length(Codigo),size.mean=mean(`Monto Estimado Adjudicado`,na.rm=T)/1000,time.mean=mean(TiempoDuracionContrato),offers.mean=mean(NumeroOferentes),totsize.per=sum(`Monto Estimado Adjudicado`,na.rm=T))%>%arrange(-n)%>%ungroup()%>%mutate(totsize.per=totsize.per/sum(totsize.per))


```

## Exploring types of project by project name

Now we turn to investigate what types of projects are developed in the sample. This is not immediate because as it was seen before there are no readily avalaible descriptive measures. We employ the name of the contract to gain some insight into what types of projects are being developed. 

First, the most common terms are usually tied to types of construction(construction, improvement, conservation) and types of work(hospitals, schools, etc.)

From the most common terms across categories we can see that project for munipalities are in the areas of social facilities(social sedes), urban development(parks, courts), and housing projects(construction, villa). In the Ministry area we have mostly health projects(hospital, cesfam) and infraestructure-related words(work). As expected, child board projects are closely related to daycare facilities, both for toddlers and infants. 


```{r projnames, echo=FALSE}
##Understand the most common words across topics
project.names=slice.projects%>%dplyr::select(Codigo,Nombre)
colnames(project.names)<-c('doc_id','text')
corpus <- DataframeSource(project.names)

corpus=VCorpus(corpus)
corpus <- tm_map(corpus, content_transformer(function(x) iconv(x = x,from = "latin1",to= "ASCII", sub="@")))
corpus <- tm_map(corpus, content_transformer(function(x) gsub('\\S+@\\S+', '', x)))

corpus <- tm_map(corpus, content_transformer(tolower))
corpus <- tm_map(corpus, content_transformer(removePunctuation))
corpus <- tm_map(corpus, content_transformer(stripWhitespace))
corpus <- tm_map(corpus, content_transformer(removeNumbers))
corpus <- tm_map(corpus, removeWords, words = c(stopwords("es")))

dtm<-DocumentTermMatrix(corpus)
freq.term=tidy(dtm)%>%group_by(term)%>%summarise(count=sum(count))%>%arrange(-count)
freq.term.merged=tidy(dtm)%>%left_join(slice.projects%>%mutate(Codigo=as.character(Codigo)),by=c('document'='Codigo'))

freq.type.table=freq.term.merged%>%group_by(term,TypeOrganism)%>%count()%>%arrange(-n)%>%ungroup()%>%group_by(TypeOrganism)%>%arrange(TypeOrganism,-n)%>%slice_head(n = 10)%>%mutate(ranking=seq_len(10))%>%pivot_wider(names_from = TypeOrganism,values_from=term,id_cols = ranking)

##Topic Modelling

total_docs=length(corpus)
k.folds<- 5
folds <- createFolds(1:total_docs, k = k.folds, list = F, returnTrain = FALSE)
candidates=c(2,3,5,8,10)
cv.results=data.frame()


for (k in candidates) {
print(k)
cv.results.intra=data.frame()
for (i in seq_len(k.folds)) {
train.corpus=corpus[folds!=i]  
test=corpus[folds==i]
dtm.train=DocumentTermMatrix(train.corpus)
raw.sum=apply(dtm.train,1,FUN=sum) #sum by raw each raw of the table
dtm.train=dtm.train[raw.sum!=0,]

dtm.test=DocumentTermMatrix(test)
raw.sum=apply(dtm.test,1,FUN=sum) #sum by raw each raw of the table
dtm.test=dtm.test[raw.sum!=0,]

lda.fitted=LDA(x = dtm.train,k)
perp.iter=topicmodels::perplexity(lda.fitted,newdata =  dtm.test)
cv.results.intra=rbind(data.frame(k=k,i=i,perplexity=perp.iter),cv.results.intra)
}
cv.results.intra=cv.results.intra%>%group_by(k)%>%summarise(mean.perp=mean(perplexity))
cv.results=rbind(cv.results.intra,cv.results)
}

ggplot(cv.results,aes(x=k,y=mean.perp))+geom_point()+geom_line()+theme_bw()+ggtitle('Perplexity by Number of topics')+xlab('Number of topics')+ylab('Mean Perplexity')

## Create a dictionary of values: construccion, etc. Tipos de proyecto. Crear un grafico de bubble con tipoconstru-tipoproyecto.


## Dimension reduction, color by unit. 


```


